---
title: "Trace, Determinant, Factorization of MatricesIS Assignment 2"
author: "Sharon Morris"
date: "2/7/2017"
output: pdf_document
---

**Problem set 1**
(1) Show that $A^T A \neq AA^T$ in general. (Proof and demonstration.)

The commutative law shows that $AB \neq BA$

```{r}
#Create matrix
A <- matrix(c(0,0,0,1,2,3,2,4,6), nrow = 3, byrow = T)
A


t(A)  # transpose the martrix

AtA <- t(A) %*% A  # t(A) times non-transposed matrix
AtA

AAt <- A %*% t(A)  #non-transposed matrix times t(A)
AAt

AtA == AAt  # Are the matrices equal
```


(2) For a special type of square matrix A, we get $A^TA = AA^T$. Under what conditions
could this be true? (Hint: The Identity matrix I is an example of such a matrix).

```{r}
A2 <- matrix(c(2,1,3,1,0,1,3,1,2), nrow = 3, byrow = T)
A2

t(A2) # transpose matrix

AA2 <- t(A2) %*% A2
AA2

A2A <- A2 %*% t(A2)
A2A

AA2 == A2A
```


**Problem set 2**
Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars.

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

You don’t have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code. 

```{r}
functA  <- function(A) {
  dimA =dim(A)[1]
  dimA
  
  idMatrix <- diag( dim(A)[1]) # crete identity matrix
  idMatrix
 
  iParameters = idMatrix   # initialize parameters
  iParameters[2,1] <- -A[2,1]/A[1,1]
  lParameters = iParameters
  
  iParameters = iParameters %*% A # inialize upper matrix
  U=lParameters
  i=1
  
  U= solve(U) # initialize lower matrix
   # loop through the column martix 
  for  (j in 1:(dimA-i)) {
    #y <- diag( dim(A)[1])
    for  (i in 3:dimA) {
      # check if pivot element is 0. 
      if (i != j) {
        if (iParameters[i,j] != 0 ) {  
          if (iParameters[j,j] < 0)  idMatrix[i,j] <- -iParameters[i,j] /     iParameters[j,j]     
          if (iParameters[j,j] > 0  )  idMatrix[i,j] <- -iParameters[i,j]/ iParameters[j,j]
          
          iParameters= idMatrix%*%iParameters # holder upper matrix
          U= U %*%solve(idMatrix) # holder lower matrix
          
          print(iParameters)
          print(U)
          idMatrix <- diag( dim(A)[1]) # reset to identity matrix
          
        }
        
      }
      
    }
  } 
  print(iParameters)
  print(U)
  U=iParameters
  s <- U %*% U
  print (s) 
  print(A)
}
```

```{r}
# Test 3x3 matrix

A <- matrix(c(2,1,0,1,2,1,0,1,2),nrow=3, byrow=TRUE)
functA(A)

# Test 4x4 matrix
A <- matrix(c(1,1,1,1,1,2,3,4,1,3,6,10,1,4,10,20),nrow=4, byrow = T)
functA(A)
```

