---
title: 'Week 1: Vectors, Matrices & Systems of Equations -- IS602'
author: "Sharon Morris"
date: "2/2/2017"
output:
  pdf_document: 
    latex_engine: lualatex
  html_document: default
---

**Problem Set 1**
(1) Calculate the dot product u.v where $u = [0.5; 0.5] and v = [3; −4]$

```{r}
u <- c(0.5,0.5)
v <- c(3,-4)
product <- sum(u*v)
product
```

(2) What are the lengths of u and v? Please note that the mathematical notion of the
length of a vector is not the same as a computer science definition.

```{r}
lenghtU <- sqrt(sum(u*u))
lenghtU

lengthV <- sqrt(sum(v*v))
lengthV
```

(3) What is the linear combination: $3u − 2v$?

```{r}
linearComb <- 3*u - 2*v
linearComb
```

(4) What is the angle between u and v?

```{r}
theta <- acos( sum(u*v) / ( sqrt(sum(u * u)) * sqrt(sum(v * v)) ) )
theta
#http://stackoverflow.com/questions/1897704/angle-between-two-vectors-in-r
```


**Problem set 2**
Set up a system of equations with 3 variables and 3 constraints and solve for x. Please write a function in R that will take two variables (matrix A & constraint vector b) and solve using elimination. Your function should produce the right answer for the system of equations for any 3-variable, 3-equation system. You don’t have to worry about degenerate cases and can safely assume that the function will only be tested with a system of equations that has a solution. Please note that you do have to worry about zero pivots, though. Please note that you should not use the built-in function solve to solve this system or use matrix inverses. The approach that you should employ is to construct an Upper Triangular Matrix and then back-substitute to get the solution. Alternatively, you can augment the matrix A with vector b and jointly apply the Gauss Jordan elimination procedure.

Please test it with the system below and it should produce a solution $x = -1.55, -0.32, 0.95$
    
    
```{r}
x = function(A, b){
  r <- dim(A)[1]
  c <- dim(A)[2]+dim(b)[2]
  upperT <- matrix(c(A, b), nrow=r, ncol=c)
  for (j in 1:(c-2)) {
    for (i in (j+1):r) {
      upperT[i,] <- upperT[i,]-upperT[j,]*upperT[i,j]/upperT[j,j]  
    }
  }
  upperT[r,] <- upperT[r,]/upperT[r,r]
  n <- numeric(r)
  n[r] = upperT[r,c]
  for (k in (r-1):1) {
    t = 0
    for (m in (k+1):r) {
      s = upperT[k,m]*n[m]
      t = t + s
    }
    n[k] = (upperT[k,c] - t) / upperT[k,k]
  }
  x <- round(n,2)
return(x)
}
```


```{r}
#Please test it with the system below and it should produce a solution x=[−1.55,−0.32,0.95]

A <- matrix(c(1, 2, -1, 1, -1, -2, 3, 5, 4), nrow=3, ncol=3) 
b <- matrix(c(1, 2, 6), nrow=3, ncol=1)
x(A,b)
```

