---
title: "Rank, Vector Spaces & Eigenvectors - IS605 Assignment 3"
author: "Sharon Morris"
date: "2/15/2017"
output:
  pdf_document: 
    latex_engine: lualatex
  html_document: default
---

**1. Problem set 1**
What is the rank of the matrix A?
$$\mathbf A = \left[\begin{array}
{rrrr}
1 & 2 & 3 & 4 \\
-1 & 0 & 1 & 3 \\
0 & 1 & -2 & 1 \\
5 & 4 & -2 & -3 \\
\end{array}\right]
$$
The rank of a matrix is the number of pivots

```{r}
library(matrixcalc)
# define the matrix
M <- matrix(data = c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), nrow = 4, ncol = 4, byrow = TRUE)
M
matrix.rank(M)

#http://stackoverflow.com/questions/10881392/rank-of-a-matrix-in-r
```

(2) Given an $$mxn$$ matrix where $$m > n$$, what can be the maximum rank? The mini- mum rank, assuming that the matrix is non-zero?

Maximum rank = $m$ (rows)
Minimum rank = 1 (all other rows could be linearly dependent)


What is the rank of matrix B?

$$\mathbf B = \left[\begin{array}
{rrr}
1 & 2 & 1 \\
3 & 6 & 3 \\
2 & 4 & 2 \\
\end{array}\right]
$$

```{r}
Mb <- matrix(data = c(1,2,1,3,6,3,2,4,2), nrow = 3, ncol = 3, byrow = TRUE)
Mb
matrix.rank(Mb)

```


**Problem set 2**
Compute the eigenvalues and eigenvectors of the matrix A. You’ll need to show your
work. You’ll need to write out the characteristic polynomial and show your solution.

$$\mathbf A = \left[\begin{array}
{rrr}
1 & 2 & 3 \\
0 & 4 & 5 \\
0 & 0 & 6 \\ 
\end{array}\right]
$$

$\lambda$ is an eigenvalue of $A$
$\iff$ $A \vec v= \lambda \vec v$ for some non zero $\vec v$
$\iff \vec 0 = \lambda I_{n}\vec v - A\vec v$ is ture
$\iff \vec 0= (\lambda I_{n}-A)\vec v$
$\iff determinant (\lambda I_{u}-A) = 0$

$$\mathbf \lambda I_{3} = \left[\begin{array}
{rrr}
\lambda & 0 & 0 \\
0 & \lambda & 0 \\
0 & 0 & \lambda \\
\end{array}\right]
$$
$$\mathbf \lambda I_{3}-A = \left[\begin{array}
{rrr}
\lambda-1 & -2 & -3 \\
0 & \lambda-4 & -5 \\
0 & 0 & \lambda-6 \\ 
\end{array}\right]
\left[\begin{array}
{rr}
\lambda-1 & -2 \\
0 & \lambda-4 \\
0 & 0  \ 
\end{array}\right]
$$

$(\lambda-1)(\lambda-4)(\lambda-6) - (\lambda -6)-(\lambda-1)-(\lambda-4)$
$=(\lambda-1)(\lambda^2-10\lambda+24)-\lambda+6-\lambda+1-\lambda+4$
$=(\lambda-1)(\lambda^2-10\lambda+24)-3\lambda+11$
$=\lambda^3-11\lambda^2+34\lambda-24-3\lambda+11$
$p(\lambda)=\lambda^3-11\lambda^2+31\lambda-13=0$
This polynomial does not have any zeros or root -- there are no possible eigenvalues




https://www.khanacademy.org/math/linear-algebra/alternate-bases/eigen-everything/v/linear-algebra-eigenvalues-of-a-3x3-matrix








