---
title: "BIAS VARIANCE TRADEOFF IN R - Assignment 12"
author: "Sharon Morris"
date: "4/17/2017"
output:
  pdf_document: default
  html_document: default
---

Using the stats and boot libraries in R perform a cross-validation experiment to observe the bias variance tradeoff. You’ll use the auto data set from previous assignments. This dataset has 392 observations across 5 variables. 

```{r}
library(stats)
library(boot)
library(ggplot2)
```

***Load Data***
```{r}
autoData <- read.table(
    "https://raw.githubusercontent.com/indianspice/IS605/master/Hw/auto-mpg.data",
    header = FALSE, as.is = TRUE) 

colnames(autoData) <- c("displacement", "horsepower","weight","acceleration","mpg")

head(autoData)
```

***Explore the Data***
```{r}
sum(is.na(autoData))
str(autoData)
```


***Polynomial Fit Model***
Fit a polynomial model of various degrees using the glm function in R and then measure the cross validation error using cv.glm function.
```{r}
n <- 8
degree <- 1:n
cv.err5 <- numeric()
for (i in degree) {
  glm.fit <- glm(mpg ~ poly(displacement + horsepower + weight + acceleration, 
                            mpg = i), data = autoData)
  cv.err5[i - min(degree) + 1] <- cv.glm(autoData, glm.fit, K = 5)$delta[1]
}
summary(glm.fit)
```


Once you have fit the various polynomials from degree 1 to 8, you can plot the cross- validation error function as
```{r}
plot(degree,cv.err5,type='b', main = "Cross Validation Estimate of Prediction Error vs. Degree")
```

The graph below shows that a degree 2 or 3 polynomial seems to fit the model the closest.


```{r}
ggplot(autoData, aes(x = poly(displacement + horsepower + weight + acceleration), 
                     y = mpg)) + geom_point()
```



**Reference**
https://www.r-bloggers.com/cross-validation-estimating-prediction-error/